Dirac neutrino mass generation from Majorana messenger

### 2019-09-25 draft.tex
Updates from Overleaf

% We will make use of the scotogenic model~\cite{Ma:2006km}. In
% scotogenic model the neutrino masses and the DM have common
% explanation. The neutrino masses are generated by one-loop effects,
% where the candidates for DM are mediators in the loop.
%Conclusion:

@@

With the aim to illustrate the one-loop Dirac neutrino mass generation
we consider the particle content shown in Table \ref{tab:partcont} as
a possible realization of the effective operator $\mathcal{O}_{6D}$. Specifically, we introduce three scalar fields
$\eta, \sigma$ and $S$, where only $S$ develops a nonzero vacuum
expectation value (VEV),  a set of three singlet fermions,
$\nu_{Rj}$ ($j=1,2$) and $\nu_{R3}$, and another set of three heavy Majorana fermions, $\psi_{R\alpha}$ ($\alpha=1,2,3$).  
%The model also includes an additional $\operatorname{U}(1)_X$ gauge
%symmetry which is used to forbid the Dirac and Majorana neutrino mass
%terms at tree level.
%Additionally, we add three right neutrinos, one scalar doublet, two scalar singlets and three Majorana fermion singlet under $\operatorname{SU}(2)_L$.
The $\operatorname{U}(1)_{\mathcal{D}}$ charges for the new particles are defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.
%In this model the neutrinos are Dirac fermions. Particle content and their respective charges are show in Tab.~\ref{tab:partcont}.

@@

%\begin{align*}
%    \mathcal{V}(H, S, \eta, \sigma) = & \sum_{i} %V(\omega_i)+\sum_{i<j}|\omega_i|^2|\omega_j|^2+ \lambda_{7} (\eta^{\dagger} H ) %(H^{\dagger} \eta) +\lambda_{8} (\eta^{\dagger} H S^{*} \sigma + \text{h.c.})\,,
%\end{align*}

@@

%Note that after the spontaneous symmetry breaking of $\operatorname{U}(1)_X$ the last term gives rise to the soft trilinear coupling
%\begin{align}
%  \mu=\lambda_8 \langle S \rangle\,,
%\end{align}

@@

% \subsection{Scalar mass spectrum}
%\label{sec:ScaMassSpect}

@@

%By minimizing the scalar potential we find that the mass-squared parameters $\mu_H^2$ and $\mu_S^2$
%\begin{align*}
%    \mu^{2}_{H} =& -\frac{1}{2}\left( \lambda_{H}  v^2 + \lambda_{1}  v_S\right)\,,  &
%    \mu^{2}_{S} =& -\frac{1}{2}\left( \lambda_{S}  v^{2}_{S} + \lambda_{1}  v^2\right)\,.
%\end{align*}

@@

%We assume that $\lambda_{1} \ll 1$, such that the scalar $S$ and $H$ do not mix. 

@@

%The other neutral mass eigenstates are defined as
%
%\begin{align*}
%    \begin{pmatrix}\chi_{(R,I)_{1}} \\ \chi_{(R,I)_{2}} \end{pmatrix} =& \begin{pmatrix} %c_{\theta_{(R,I)}} & -s_{\theta_{(R,I)}} \\ s_{\theta_{(R,I)}} & c_{\theta_{(R,I)}} \end{pmatrix} %\begin{pmatrix}\sigma_{(R,I)} \\ \eta_{(R,I)} \end{pmatrix} \,,
%\end{align*}
%
%where $\tan2\theta_{(R,I)} = \pm2c/(b-a)$, with $a = m^{2}_{\eta^{\pm}} + \frac{1}{2}\lambda_{7} %\upsilon^{2}_{H}$, $b = \mu^{2}_{\sigma}+\frac{1}{2}(\lambda_{2} \upsilon^{2}_{H} + \lambda_{4} %\upsilon^{2}_{S})$ and $c = \frac{1}{2}\lambda_8 v_{H} v_{S}$\,. 
%The corresponding mass eigenvalues are
%\begin{align*}
%   m_{\chi_{R(1,2)}}^{2} &= \frac{1}{2} \left(a + b \mp \sqrt{(a-b)^{2} + 4c^{2}} \right)\,, \\
%     m_{\chi_{I(1,2)}}^{2} &= \frac{1}{2} \left(a + b \mp \sqrt{(a-b)^{2} - 4c^{2}} \right)\,,
%\end{align*}

On the other hand, we assume that the heavy Majorana fermions are already in the diagonal basis in such a way their masses are $M_{\psi_\alpha}=\kappa_{\alpha\alpha}v_S/\sqrt{2}$, with $M_{\psi_1}<M_{\psi_2}<M_{\psi_3}$. Finally, the mass of the new gauge boson is given by 
\begin{align}
    M_{Z'}=q_Sg^{\prime} v_S.
\end{align}
%\begin{pmatrix}
  %\psi_1 &
  %\psi_2 &
  %\psi_3
%\end{pmatrix}^{\operatorname{T}}$$

\section{Neutrino masses and charged lepton flavor violation}
\label{sec:Neutrinos}
%
\begin{figure}
\centering
\includegraphics[scale=0.6]{Neutrino_Loop.pdf}
\caption{One-loop realization of the dimension-6 operator $\overline{L} \tilde{H} \nu_R S^2$ leading to Dirac neutrino masses with Majorana mediators.}
\label{fig:zee}
\end{figure}
%
%Thanks to the scalar mixing term $\lambda_8$ nonzero Dirac
Neutrino masses are generated at one-loop level according to the diagram in  Fig.~\ref{fig:zee}. The expression for the effective neutrino mass matrix $\mathcal{M}_{\nu}$ can be cast as
%
\begin{align}
(\mathcal{M}_{\nu})_{ij} = \frac{1}{32 \pi^{2}}  \frac{\lambda_8 v_S v_H} {m_{\chi_{R_2}}^{2}-m_{\chi_{R_1}}^{2}}\sum_{\alpha=1}^{3} h_{i \alpha} M_{\psi_\alpha} y^{*}_{j\alpha}\left[ F\left( \frac{m_{\chi_{R_2}}^{2}}{M_{\psi_{\alpha}}^{2}} \right) - F\left( \frac{m_{\chi_{R_1}}^{2}}{M_{\psi_{\alpha}}^{2}} \right) \right] + (R \to I)\,,
%(\mathcal{M}_{\nu})_{ij} = \frac{1}{64 \pi^{2}}  \frac{\lambda_8 v_S v_H} %{m_{\chi_{R_2}}^{2}-m_{\chi_{R_1}}^{2}}\sum_{\alpha=1}^{3} h_{i \alpha} %M_{\psi_{\alpha}}y^{*}_{j\alpha}\left[ F\left( %\frac{m_{\chi_{R_2}}^{2}}{M_{\psi_{\alpha}}^{2}} \right) - F\left( %\frac{m_{\chi_{R_1}}^{2}}{M_{\psi_{\alpha}}^{2}} \right) \right] + (R \to I)\,,
\end{align}
%
%check the following formula:
where $F(x) =x \log x/(x-1)$. 
%, $M_{\psi_{a}}$ is
%the fermion mass and $\chi_{R_{(1,2)}}$ ( $\chi_{I_{(1,2)}}$ ) are the
%two CP-even (CP-odd) mass eigenstates.
%It follows that to obtain nonzero neutrino masses it is necessary that $\mu\equiv\lambda_8 %v_S \neq 0$. 
Note that the structure of the effective neutrino mass matrix, given by the
product $(M_{\nu})_{ij} \propto h_{i \alpha} y_{j\alpha}$, is similar to
the structure of the neutrino mass matrix for the tree-level seesaw
mechanism for Dirac neutrinos~\cite{Chulia:2016ngi}. 
It is also worth mentioning that if only one
fermion $\psi_R$ is added, then there will be two massless neutrinos, which
would be ruled out by the current neutrino oscillation data~\cite{deSalas:2017kay}. 
In our case, we assume the
existence of three of such fermions, generating Dirac scotogenic masses for
the two left-handed neutrinos ($\nu_{3}=\nu_{L3}+\nu_{R3}$ is massless due to the charge assignment).

@@

%The mass matrix for Dirac neutrinos is diagonalized by
%\begin{align*}
%(U^{\nu}_R)^{\dagger} \mathcal{M}_{\nu} U^{\nu}_L = \mathcal{M}_{\nu}^{\text{diag}} = %\text{diag}(m_1, m_2, m_3)\,,
%\end{align*}
%where $U^{(\nu)}_R$ and $U^{(\nu)}_L$ are the unitary $3 \times 3$
%matrices associated to the right-handed and left-handed neutrino sectors. The leptonic %charged currents with non-massless neutrinos
%are parametrized in terms of the unitary mass
%\begin{align*}
%U_{\text{PMNS}} = \left( U^{\ell}_L \right)^{\dagger} U^{\nu}_L\,,
%\end{align*}
%where is the left diagonalization matrix of the charged leptons.
%As we already are in the diagonal basis after the calculation of the
%one-loop neutrino masses, we can set
%$U_{L}^{\ell} = \mathds{1}$. Moreover, we assume that $U_{R}^{\nu} = \mathds{1}$.
%Therefore, we can express the Dirac neutrino mass matrix in the
%following way
%\begin{align}
%(\mathcal{M}_{\nu})_{ij} = (U_{\text{PMNS}})_{ij} (\mathcal{M}_{\nu}^{\text{diag}})_{j}\,,
%\end{align}
%where the matrix $U_{\text{PMNS}}$ is expressed in the standard parameterization as
%\begin{align*}
%U_{\text{PMNS}} = \begin{pmatrix}
%    1 & 0		& 0 \\
%    0 & c_{23}  & s_{23} \\
%    0 & -s_{23} & c_{23}
%\end{pmatrix}
%\begin{pmatrix}
%    c_{13}  &  0 & s_{13}e^{-i\delta} \\
%    0 		&  1 & 0 \\
%    -s_{13}e^{i\delta} &  0 & c_{13}
%\end{pmatrix}
%\begin{pmatrix}
%    c_{12}  & s_{12} & 0 \\
%    -s_{12} & c_{12} & 0 \\
%    0		& 0		 & 1
%\end{pmatrix}\,,
%\end{align*}
%where $s_{ij} = \sin \theta_{ij}$ and $c_{ij} = \cos \theta_{ij}$ and $\theta_{ij}$ are %neutrino mixing angles and $\delta$ is CP-violating phase.

%\section{Lepton flavor violation}
%\label{sec:LFV}

@@

%
%which leads to a very slight requirement,

%The most common search for CLFV focuses on the $l^{-}_{i} \to l^{-}_{j} \gamma$ process. This process is described by the effective Lagrangian
%
%\begin{align}
%    \mathcal{L} = \frac{\mu_{m}^{\alpha \beta}}{2} l^{-}_{\beta} \sigma^{\mu \nu} l^{-}_{\alpha} F_{\mu \nu}
%\end{align}
%
%where $\mu_{m}^{\alpha \beta}$ is a transition magnetic moment~\cite{Toma:2013zsa}.
%The decay rate is given by~\cite{Lavoura:2003xp}
%
%\begin{align}
% \Gamma(l^{-}_{i} \to l^{-}_{j} \gamma) = \frac{m^{3}_{i}}{16 \pi} \left| \sum_{\alpha} h_{i \alpha} m_{i} h_{\alpha j} \frac{i e}{16 \pi M^{2}_{\eta^{-}}} K(t_{k}) \right|^{2}\,, 
%\end{align}
%
%where $t_{\alpha} = M^{2}_{\psi_{\alpha}}/M^{2}_{\eta^{-}}$ and
%
%\begin{align}
%    K(t_{\alpha}) = \frac{2t_{\alpha}^{2}+5t_{\alpha}-1}{12(t_{\alpha}-1)^{3}} - %\frac{t_{\alpha}^{2}\log t_{\alpha}}{2(t_{\alpha}-1)^{4}}\,.
%\end{align}
%
%If we assume that $t_{\alpha} \to 0$, $t_{\alpha} \to 1$ and $t_{\alpha} \to \infty$, we have that $K(t_{\alpha}) = 1/12$, $K(t_{\alpha}) = 1/24$ and $K(t_{\alpha}) = 1/(6t)$, respectively.

%For $i \neq j$ we have CLFV processes, these processes are severely restricted by the current bounds , which set limits to the Yukawas. The most restrictive experimental bounds is decay $\mu \to e \gamma$, which is several orders of magnitude greater than radiative decays $\tau \to~l\gamma$, with $l = e, \mu$. These restrictions are given by Br$(\mu \to e\gamma)$ \textless $5.7 \times 10^{-13} $~\cite{Adam:2013mnn}, Br$(\tau \to e\gamma) $\textless$ 3.3 \times 10^{-8}$ and Br$(\tau \to~\mu\gamma) $\textless$ 4.4 \times 10^{-8}$~\cite{Aubert:2009ag, Bona:2007qt, Miyazaki:2012mx}.

%In the limit $m_{\eta^{\pm}}^{2} \gg M_{\psi_{a}}^{2}$, $t_{\alpha} \to 0$ the decay width reads
%
%\begin{align}
% \Gamma(l^{-}_{i} \to l^{-}_{j} \gamma) = \frac{m^{3}_{i}}{16 \pi} \left| \sum_{\alpha} h_{i \alpha} m_{i} h_{\alpha j} \frac{i e}{16 \pi M^{2}_{\eta^{-}}} \left[ \frac{1}{12} \right] \right|^{2}\,, 
%\end{align}
%
%using the current experimental constraint to Br$(\mu \to e\gamma)$, can be fullfilled provided
%
%\begin{align*}
%    \left( \sum_{\alpha} \frac{h^{}_{i \alpha} h_{\alpha j}^{*}}{M^{2}_{\eta^{-}}} \right)^{2} \leq 5.7 \times 10^{-13} \frac{768 \pi G_{F}^{2}}{\alpha_{EM}}
%\end{align*}
%
%which leads to a very slight requirement,
%\begin{align*}
%    \left| \sum_{\alpha}h^{}_{i \alpha} h_{\alpha j}^{*}\left(\frac{50\text{TeV}}{M^{2}_{\eta^{-}}} \right)^{2} \right| \leq 12.65
%\end{align*}
%
%For $i = j$ we have processes that contribute to the leptonic magnetic dipole moment. In the case of the anomalous muon magnetic dipole moment, the experimental measurements does not agree with the values predicted by the SM~\cite{Lindner:2016bgg}. This can be a proof of the veracity of the model, in addition to giving restrictions on the Yukawas.

@@

%\subsection{Fermion dark matter}
Since $\psi_{1}$ is a singlet under the SM gauge group its thermal relic density is controlled by the Yukawa couplings and $\operatorname{U}(1)_{\mathcal{D}}$ gauge interactions. 
The scenario where $\psi_{1}$ self-annihilates dominantly through $h_{i\alpha}$-mediated interactions resembles the very well known scotogenic model \cite{Ma:2006km},  
where sizable $h_{i\alpha}$ are required to reproduce the correct DM abundance, which in turn leads to a mild tension with experimental upper bounds on the rates for rare charged lepton decays~\cite{Kubo:2006yx,Sierra:2008wj,Ibarra:2016dlb}. 
%The phenomenology of this scenario has been extensively studied  \cite{Kubo:2006yx,Sierra:2008wj,Suematsu:2009ww,Gelmini:2009xd,Hambye:2009pw,Aoki:2010tf,Schmidt:2012yg,Ma:2012if,Kashiwase:2012xd,Kashiwase:2013uy,Racker:2013lua,Toma:2013zsa,Molinaro:2014lfa,Vicente:2014wga,Merle:2015gea,Merle:2015ica,Merle:2016scw,Hagedorn:2016dze,Lindner:2016kqk,Hessler:2016kwm,Ibarra:2016dlb}.
On the other hand, when the $Z'$ portal \cite{Langacker:2008yv,Arcadi:2013qia} is the main gate to visible sector, $\psi_1$ largely annihilates into neutrinos in such a way the observed DM abundance can be reproduced without entering in conflict with the DM searches, which follows from the fact that the $Z'$ does not couple to quarks and charged leptons
(see~\cite{Arcadi:2013qia,Alves:2013tqa,Alves:2015pea,Alves:2016cqf,Blanco:2019hah} for phenomenological studies on $Z'$-mediated Majorana DM). 
% check Ibarra papers
%also through the annihilation
%$\psi_{1} \psi_{1} \to l _{\alpha}^+ \overline{l}_{\beta}^- $, via 
%Yukawa coupling $h^{a}_{i}$. These Yukawa couplings are restricted by
%CLFV and neutrino masses. %But, Yukawa couplings $y^{a}_{j}$ is free.
%In the case where the mass of the other fermions and scalars is nearly
%degenerated to $\psi_{1} $, the processes of coannihilation can be
%relevant for the relic density calculation~\cite{Klasen:2013jpa}.

%   Because the particle $\psi_{1}$ only interacts with the leptons via the Yukawa couplings, this model has no a direct detection cross section at three level.

% In the case that the mass of the lightest fermion $\psi$ is smaller
% than the mass of the scalars, the DM will be fermionic. For this DM it
% is possible to have the following observable decay of the scalar
% $\eta^{\pm} \to l^{\pm} \psi_{1}$, therefore, the other fermions decay
% as $ \psi \to l^{\pm} l^{\mp} \psi_{1}$ and
% $\psi \to l^{\pm} l^{\mp} \psi_{1,2} $, through $\eta^{\pm}$.

%\subsection{Scalar dark matter}
In contrast to fermion DM, the scalar DM candidate has additional interaction terms to the Yukawa and gauge interactions. This entails that the later ones can be used to alter the relic density predictions in scenarios with mixed scalar DM. Since in the present model the CP-even and CP-odd neutral $Z_2$-odd particles are mass degenerate, we have the scenario of singlet-doublet complex DM~\cite{Kadastik:2009dj,Belanger:2012vp} where the DM candidate is a mixture of a complex singlet \cite{McDonald:1993ex} and a $SU(2)_L$ doublet \cite{Deshpande:1977rw,Barbieri:2006dq}. It follows that for negligible Yukawa and gauge interactions there are two DM mass regions that allow us to properly reproduce the relic abundance, one corresponds the Higgs funnel region and the second one demands masses above $100$ GeV~\cite{Kakizaki:2016dza}.

%This corresponds to the simplified dark matter model of scalar singlet-doublet dark matter, which have been studied in detail in~\cite{Restrepo:2019soi}

%When the lightest scalar  is mainly a scalar doublet, the phenomenology will be similar inert doublet model~\cite{Honorez:2010re}. For this kind of DM model  inert doublet model there are two mass regions for DM which can to reproduce adequately the relic density ($50$GeV \textless $m_{DM}$ \textless $70$GeV and $535$GeV \textless $m_{DM}$ \textless $20$TeV)~\cite{Garcia-Cely:2015khw}.

% In recent studies, these two regions are connected through use of a
% second doublet~\cite{Borah:2019aeq}. How we have in the model the
% Yukawa coupling, the decay $\psi_{i} \to l^{\pm} \eta^{\mp}$ is
% possible. In addition, decay is given via scalar interactions
% $\eta^{\pm} \to \eta^{0} + {W^{\pm}}$, where $W^{\pm}$ can be real or
% virtual and decay to leptons or quarks.

%Now, if the lightest scalar is mainly a singlet, the DM will be similar to Singlet %Scalar Dark Matter Model~\cite{Athron:2017kgt}\todo{change for the original ref by Ma et al. etc}.
%The mass range for the scalar singlet that produces the adequate relic
%density is around the Higgs resonance and for masses above $1$TeV. For
%this model~\cite{}, the relevant coupling for DM phenomenology is the
%Higgs portal trough the Lagrangian term $\sigma^{2} H^{2}_{i}$.

%This term generates interesting phenomenological consequences.
%Such as, thermal production of DM in the early
%universe~\cite{Yaguna:2008hd}, direct detection and invisible decay of
%SM Higgs through $h \to \sigma \sigma$~\cite{Mambrini:2011ik}.
%In addition, this model has important contributions for cosmology, in
%phenomena such as inflation~\cite{Lerner:2009xg} and
%baryogenesis~\cite{Cline:2012hg}.

@@

With the aim of thermalize the right-handed neutrinos with the
primordial plasma in the early Universe, we consider an modification of
the previous setup by considering a general anomaly free Abelian gauge
symmetry with generation-independent charge assignments for the
SM fermions, $\operatorname{U}(1)_{X}$.
% 
In the Appendix we show that the solutions to the anomaly cancellation conditions
allow us to write the $X$-charges of the SM fields in terms
of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc,Allanach:2018vjg}, that we choose as
\begin{align}
  X(r,l)=r R - l Y\,,
\end{align}
where $R$ is the generator of $\operatorname{U}(1)_R$ (the gauge Abelian symmetry where only the right-handed SM fermions have non-vanishing $X$-charges), $Y$ is the hypercharge, $l$ is the $X$-charge of the lepton doublets, and $r$ parametrizes the contribution to the linear and mixed gauge-gravitational anomalies of any extra set of chiral fermions, as given in Eq.~\eqref{eq:anolam} of the Appendix.
In this way, if only extra vector-like fermions are allowed beyond the SM ($r=0$) the solution must be proportional to hypercharge.
%
Without lost of generality, we can write the solutions in terms of just one parameter~\cite{Jenkins:1987ue,Oda:2015gna,Okada:2018tgy}, that we choose to be $l$  after fix $r=1$. Then, as shown in the Appendix, the full family of solutions for a fixed $l$ can be obtained after  rescaling all the $X$-charges by a factor $r$. In particular, the fermiophobic~\cite{Campos:2017dgc,Bertuzzo:2018itn,Bertuzzo:2018ftf} solution used in the previous sections, $\operatorname{U}(1)_{\mathcal{D}}$, corresponds to the rescaling $r=0$ of the $\operatorname{U}(1)_R$ solution: $\mathcal{D}=X(0,0)$.

@@

%Note that if $r$ is a fraction with denominator $4$, let say $r=3/4$, 
%The $\operatorname{U}(1)_{B-L}$ left out a remnant $Z_2$ discrete symmetry which guarantees the stability of the lightest odd-particle.

%the charges for the SM particles are in equations~\ref{Eq:SMCharges}. Since these are determined by the anomaly cancellation equations.
%For this operator we have that the charges for the new particles are given by
%\begin{align}
%    \psi=&-\frac{\nu+1}{4}\,,&\eta=&-\frac{\nu+1}{4}-l\,,&\sigma=&\frac{1-3\nu}{4}\,.
%\end{align}
%As we are interested here in the case of the exotic set of X-charges charges:$\mp 4$,$\mp 4$,$\pm 5$, the charges for the model particles are shown in the table~\ref{tab:partcont3}.
%Because the charges of the SM fermions are the same for both operators ($5d$ and $6d$), the $N_{eff}$ restrictions shown in the figure~\ref{fig:Neff} apply to both operators.
%
\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c||c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ &  $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_R$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$ & $\operatorname{U}(1)_{\mathcal{D}}$\\ \hline
$L $     & $\boldsymbol{2}$ & $-1$ & $l$      &  $-1$&    $0$ &  $-3/2$&  $-1/2$ & $0$ \\    
$d_R $   & $\boldsymbol{1}$ & $-2/3$ & $1+2l/3$ &  $1/3$&    $1$&  $0$&    $2/3$ & $0$\\
$u_R $   & $\boldsymbol{1}$ & $+4/3$ & $-1-4l/3$&  $1/3$&   $-1$&  $1$&   $-1/3$ & $0$ \\
$Q $     & $\boldsymbol{2}$ & $1/3$ & $-l/3$   & $1/3$&    $0$&  $1/2$&  $1/6$ & $0$ \\
$e_R $   & $\boldsymbol{1}$ & $-2$   & $1+2l$   &  $-1$&    $1$ &  $-2$&  $0$ & $0$\\\hline
$H $     & $\boldsymbol{2}$ & $1$  & $-1-l$   &  $0$&    $-1$ &  $1/2$&  $-1/2$ & $0$ \\
$\eta$   & $\boldsymbol{2}$ & $1$ & $3/4-l$  & $7/4$& $3/4$ &$9/4$&$5/4$ & $1$ \\
$S$      & $\boldsymbol{1}$ & $0$    & $3/2$    & $3/2$&  $3/2$ & $3/2$& $3/2$ & $2$\\
$\sigma$ & $\boldsymbol{1}$ & $0$    & $13/4$   & $13/4$&  $13/4$ & $13/4$& $13/4$ & $3$\\\hline
$\nu_{Ri}$& $\boldsymbol{1}$ & $0$   & $-4$& $-4$&  $-4$ & $-4$& $-4$ & $-4$\\
$\nu_{R3}$& $\boldsymbol{1}$ & $0$   & $+5$& $+5$&  $5$ & $5$& $5$ & $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/4$ & $3/4$&  $3/4$ & $3/4$& $3/4$ & $1$ \\\hline
$\xi_{L\alpha}$   & $\boldsymbol{1}$ & $0$ & $3/4$& $3/4$ &  $3/4$ & $3/4$& $3/4$ &$-$\\\hline
  \end{tabular}
  \caption{General one-parameter solution with some examples of rational solutions
    ($X=B-L,R,D,G$ and $\mathcal{D}$) for the radiative type-I seesaw realization
    of the effective operator $\mathcal{O}_{6D}$ for Dirac neutrino masses. The
    last column  corresponds to the solution in Table~\ref{tab:partcont}.}
    \label{tab:partcont3}
\end{table}
%

% \footnote{Check 2r+2nu: -2,6,7,-9,-3  3r+2nu: -(5/2),7/2,-7,6,9/2}

Accordingly, we may expect that 
the three right-handed neutrinos within the not-so minimal model
contribute to the radiation energy density of the Universe, since now   the interaction between the $Z'$ with the SM fermions  opens up the possibility to thermalize them with the primordial plasma.   
%These new contributions can modify the cosmological observables in the cosmic microwave background (CMB). 
%These new neutrinos contribute to the
%existence of dark radiation, which affects the expansion processes of

>>>>

@@

With the aim to illustrate the one-loop Dirac neutrino mass generation
we consider the particle content shown in Table \ref{tab:partcont} as
a possible realization of the effective operator $\mathcal{O}_{6D}$. Specifically, we introduce three scalar fields
$\eta, \sigma$ and $S$, where only $S$ develops a nonzero vacuum
expectation value (VEV),  a set of three singlet fermions,
$\nu_{Rj}$ ($j=1,2$) and $\nu_{R3}$, and another set of three heavy Majorana fermions, $\psi_{R\alpha}$ ($\alpha=1,2,3$).  
The $\operatorname{U}(1)_{\mathcal{D}}$ charges for the new particles are defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.

@@

@@

@@

@@

@@

@@

On the other hand, we assume that the heavy Majorana fermions are already in the diagonal basis in such a way their masses are $M_{\psi_\alpha}=\kappa_{\alpha\alpha}v_S/\sqrt{2}$, with $M_{\psi_1}<M_{\psi_2}<M_{\psi_3}$. Finally, the mass of the new gauge boson is given by 
\begin{align}
    M_{Z'}=q_Sg^{\prime} v_S.
\end{align}

\section{Neutrino masses and charged lepton flavor violation}
\label{sec:Neutrinos}
%
\begin{figure}
\centering
\includegraphics[scale=0.6]{Neutrino_Loop.pdf}
\caption{One-loop realization of the dimension-6 operator $\overline{L} \tilde{H} \nu_R S^2$ leading to Dirac neutrino masses with Majorana mediators.}
\label{fig:zee}
\end{figure}
%
Neutrino masses are generated at one-loop level according to the diagram in  Fig.~\ref{fig:zee}. The expression for the effective neutrino mass matrix $\mathcal{M}_{\nu}$ can be cast as
%
\begin{align}
(\mathcal{M}_{\nu})_{ij} = \frac{1}{32 \pi^{2}}  \frac{\lambda_8 v_S v_H} {m_{\chi_{R_2}}^{2}-m_{\chi_{R_1}}^{2}}\sum_{\alpha=1}^{3} h_{i \alpha} M_{\psi_\alpha} y^{*}_{j\alpha}\left[ F\left( \frac{m_{\chi_{R_2}}^{2}}{M_{\psi_{\alpha}}^{2}} \right) - F\left( \frac{m_{\chi_{R_1}}^{2}}{M_{\psi_{\alpha}}^{2}} \right) \right] + (R \to I)\,,
\end{align}
%
where $F(x) =x \log x/(x-1)$. 
Note that the structure of the effective neutrino mass matrix, given by the
product $(M_{\nu})_{ij} \propto h_{i \alpha} y_{j\alpha}$, is similar to
the structure of the neutrino mass matrix for the tree-level seesaw
mechanism for Dirac neutrinos~\cite{Chulia:2016ngi}. 
It is also worth mentioning that if only one
fermion $\psi_R$ is added, then there will be two massless neutrinos, which
would be ruled out by the current neutrino oscillation data~\cite{deSalas:2017kay}. 
In our case, we assume the
existence of three of such fermions, generating Dirac scotogenic masses for
the two left-handed neutrinos ($\nu_{3}=\nu_{L3}+\nu_{R3}$ is massless due to the charge assignment).

@@

@@

@@

Since $\psi_{1}$ is a singlet under the SM gauge group its thermal relic density is controlled by the Yukawa couplings and $\operatorname{U}(1)_{\mathcal{D}}$ gauge interactions. 
The scenario where $\psi_{1}$ self-annihilates dominantly through $h_{i\alpha}$-mediated interactions resembles the very well known scotogenic model \cite{Ma:2006km},  
where sizable $h_{i\alpha}$ are required to reproduce the correct DM abundance, which in turn leads to a mild tension with experimental upper bounds on the rates for rare charged lepton decays~\cite{Kubo:2006yx,Sierra:2008wj,Ibarra:2016dlb}. 
On the other hand, when the $Z'$ portal \cite{Langacker:2008yv,Arcadi:2013qia} is the main gate to visible sector, $\psi_1$ largely annihilates into neutrinos in such a way the observed DM abundance can be reproduced without entering in conflict with the DM searches, which follows from the fact that the $Z'$ does not couple to quarks and charged leptons
(see~\cite{Arcadi:2013qia,Alves:2013tqa,Alves:2015pea,Alves:2016cqf,Blanco:2019hah} for phenomenological studies on $Z'$-mediated Majorana DM).

In contrast to fermion DM, the scalar DM candidate has additional interaction terms to the Yukawa and gauge interactions. This entails that the later ones can be used to alter the relic density predictions in scenarios with mixed scalar DM. Since in the present model the CP-even and CP-odd neutral $Z_2$-odd particles are mass degenerate, we have the scenario of singlet-doublet complex DM~\cite{Kadastik:2009dj,Belanger:2012vp} where the DM candidate is a mixture of a complex singlet \cite{McDonald:1993ex} and a $SU(2)_L$ doublet \cite{Deshpande:1977rw,Barbieri:2006dq}. It follows that for negligible Yukawa and gauge interactions there are two DM mass regions that allow us to properly reproduce the relic abundance, one corresponds the Higgs funnel region and the second one demands masses above $100$ GeV~\cite{Kakizaki:2016dza}.

@@

With the aim of thermalize the right-handed neutrinos with the
primordial plasma in the early Universe, we consider an modification of
the previous setup by considering a general anomaly free Abelian gauge
symmetry with generation-independent charge assignments for the
SM fermions, $\operatorname{U}(1)_{X}$.
% 
In the Appendix we show that the solutions to the anomaly cancellation conditions
allow us to write the $X$-charges of the SM fields in terms
of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc,Allanach:2018vjg,Das:2019pua}, that we choose as
\begin{align}
  X(r,l)=r R - l Y\,,
\end{align}
where $R$ is the generator of $\operatorname{U}(1)_R$ (the gauge Abelian symmetry where only the right-handed SM fermions have non-vanishing $X$-charges), $Y$ is the hypercharge, $l$ is the $X$-charge of the lepton doublets, and $r$ parametrizes the contribution to the linear and mixed gauge-gravitational anomalies of any extra set of chiral fermions, as given in Eq.~\eqref{eq:anolam} of the Appendix.
In this way, if only extra vector-like fermions are allowed beyond the SM ($r=0$) the solution must be proportional to hypercharge.
%
Without lost of generality, we can write the solutions in terms of just one parameter~\cite{Jenkins:1987ue,Oda:2015gna,Okada:2018tgy}, that we choose to be $l$  after fix $r=1$. Then, as shown in the Appendix, the full family of solutions for a fixed $l$ can be obtained after  rescaling all the $X$-charges by a factor $r$. In particular, the fermiophobic~\cite{Campos:2017dgc,Bertuzzo:2018itn,Bertuzzo:2018ftf} solution used in the previous sections, $\operatorname{U}(1)_{\mathcal{D}}$, corresponds to the rescaling $r=0$ of the $\operatorname{U}(1)_R$ solution: $\mathcal{D}=X(0,0)$.

@@

%
\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c||c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ &  $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_R$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$ & $\operatorname{U}(1)_{\mathcal{D}}$\\ \hline
$L $     & $\boldsymbol{2}$ & $-1$ & $l$      &  $-1$&    $0$ &  $-3/2$&  $-1/2$ & $0$ \\    
$d_R $   & $\boldsymbol{1}$ & $-2/3$ & $1+2l/3$ &  $1/3$&    $1$&  $0$&    $2/3$ & $0$\\
$u_R $   & $\boldsymbol{1}$ & $+4/3$ & $-1-4l/3$&  $1/3$&   $-1$&  $1$&   $-1/3$ & $0$ \\
$Q $     & $\boldsymbol{2}$ & $1/3$ & $-l/3$   & $1/3$&    $0$&  $1/2$&  $1/6$ & $0$ \\
$e_R $   & $\boldsymbol{1}$ & $-2$   & $1+2l$   &  $-1$&    $1$ &  $-2$&  $0$ & $0$\\\hline
$H $     & $\boldsymbol{2}$ & $1$  & $-1-l$   &  $0$&    $-1$ &  $1/2$&  $-1/2$ & $0$ \\
$\eta$   & $\boldsymbol{2}$ & $1$ & $3/4-l$  & $7/4$& $3/4$ &$9/4$&$5/4$ & $1$ \\
$S$      & $\boldsymbol{1}$ & $0$    & $3/2$    & $3/2$&  $3/2$ & $3/2$& $3/2$ & $2$\\
$\sigma$ & $\boldsymbol{1}$ & $0$    & $13/4$   & $13/4$&  $13/4$ & $13/4$& $13/4$ & $3$\\\hline
$\nu_{Ri}$& $\boldsymbol{1}$ & $0$   & $-4$& $-4$&  $-4$ & $-4$& $-4$ & $-4$\\
$\nu_{R3}$& $\boldsymbol{1}$ & $0$   & $+5$& $+5$&  $5$ & $5$& $5$ & $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/4$ & $3/4$&  $3/4$ & $3/4$& $3/4$ & $1$ \\\hline
$\xi_{L\alpha}$   & $\boldsymbol{1}$ & $0$ & $3/4$& $3/4$ &  $3/4$ & $3/4$& $3/4$ &$-$\\\hline
  \end{tabular}
  \caption{General one-parameter solution with some examples of rational solutions
    ($X=B-L,R,D,G$ and $\mathcal{D}$) for the radiative type-I seesaw realization
    of the effective operator $\mathcal{O}_{6D}$ for Dirac neutrino masses. The
    last column  corresponds to the solution in Table~\ref{tab:partcont}.}
    \label{tab:partcont3}
\end{table}
%
